---
title: "MATH381-HW7"
author: "Ethan Zhuang"
date: "3/1/2022"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(RColorBrewer)
library(wordcloud)
```
## R Markdown

The code reads in the 8-dimensional vector
information from the file “hartigan-file22.csv” into a table called
“cities”, which has the object name, the cities

```{r}
cities <- read.csv("/Users/yuhaozhuang/Desktop/MATH381/HW7/hartigan-file22.csv",header=FALSE)
cities <- subset(cities,select=-V2)
cities
```

The code gives a matrix of expection values without the name of object, cities.
```{r}
cities2 <- subset(cities,select=-V1)
cities2 <- cities2[-c(6), ] 
cities2
```

This command takes the table “cities2” and creates a matrix of
distances between every row of the table after applying the
function “myfun” to every in the table (MARGIN=c(1,2)
specifies over all rows and columns).
```{r}
d <- dist(cities2, method="minkowski",p=2)
d
```

```{r}

library(BBmisc)
```
Nomalized the data by finding the mean and standard deviation of each column, and then subtracting the mean and divide by the standard deviation in each column. And creates the corresponding distances table.
```{r}
# Name this standardization method as normalize1.
normalize1 <- normalize(cities2,method="standardize",margin=2)
d1 = dist(normalize1, method="minkowski", p = 2)
d1
```

Normalized the data by finding the min and max value in each column, and then mapping each column entry x to (x-min)/(max-min). And creates the corresponding distances table.
```{r}
# Name this standardization method as normalize2.
normalize2 <- matrix(nrow = length(cities2[, 1]), ncol = length(cities2))
for (j in 1:length(cities2)) {
  for (i in 1:length(cities2[, j])) {
    normalize2[i, j] <- (cities2[i, j] - min(cities2[, j]))/(max(cities2[, j]) - min(cities2[, j]))
  }
}
d2 = dist(normalize2, method="minkowski", p = 2)
d2
```
## Models for non-normalization data

```{r}
# 1-D model
model1_1 <- cmdscale(d,k=1,eig=TRUE)
# points in the model
points1_1 <- model1_1$points
points1_1
```

```{r}
# eigenvalues in the model
eigs1_1 <- model1_1$eig
eigs1_1
```

```{r}
# Plot eigenvalues
plot(model1_1$eig, xlab = "Index for cities", ylab = "Eigenvalues")
```

```{r}
# GOF of the 1-D model
GOF1_1 <- model1_1$GOF
GOF1_1
```

```{r}
library(ggrepel)
# labels
cityIndex <- 1:length(cities2[, 1])
# Visulize 1-D model
m1_1 = data.frame(model1_1$points, 0)
textplot(m1_1[, 1], m1_1[, 2], gsub("(ˆ[ˆ\\s]+\\s{1})", "", cityIndex, perl = TRUE), xlim = c(-10, 10), ylim = c(-10, 10), cex = 0.6)
```

```{r}
# the distances in the model compare with the original distances
plot(d, dist(points1_1))
abline(0, 1, col = "red")
```

```{r}
# mean absolute difference of the entries
diff1_1a <- mean(abs(as.matrix(dist(points1_1)) - as.matrix(d)))
diff1_1a
```
```{r}
# maximum absolute difference of the entries
diff1_1b <- max(abs(as.matrix(dist(points1_1)) - as.matrix(d)))
diff1_1b
```

```{r}
# 2-D model
model1_2 <- cmdscale(d,k=2,eig=TRUE)
# points in the model
points1_2 <- model1_2$points
points1_2
```

```{r}
# eigenvalues in the model
eigs1_2 <- model1_2$eig
eigs1_2
```

```{r}
# Plot eigenvalues
plot(model1_2$eig, xlab = "Index for cities", ylab = "Eigenvalues")
```

```{r}
# GOF of the 2-D model
GOF1_2 <- model1_2$GOF
GOF1_2
```

```{r}
library(ggrepel)
# labels
cityIndex <- 1:length(cities2[, 1])
# Visulize 2-D model
textplot(points1_2[, 1], points1_2[, 2], gsub("(ˆ[ˆ\\s]+\\s{1})", "", cityIndex, perl = TRUE), xlim = c(-10, 40), ylim = c(-10, 40), cex = 0.6)
```

```{r}
# the distances in the model compare with the original distances
plot(d, dist(points1_2))
abline(0, 1, col = "red")
```

```{r}
# mean absolute difference of the entries
diff1_2a <- mean(abs(as.matrix(dist(points1_2)) - as.matrix(d)))
diff1_2a
```
```{r}
# max absolute difference of the entries
diff1_2b <- max(abs(as.matrix(dist(points1_2)) - as.matrix(d)))
diff1_2b
```

```{r}
# 3-D model
model1_3 <- cmdscale(d,k=3,eig=TRUE)
# points in the model
points1_3 <- model1_3$points
points1_3
```

```{r}
# eigenvalues in the model
eigs1_3 <- model1_3$eig
eigs1_3
```

```{r}
# Plot eigenvalues
plot(model1_3$eig, xlab = "Index for cities", ylab = "Eigenvalues")
```

```{r}
# GOF of the 3-D model
GOF1_3 <- model1_3$GOF
GOF1_3
```

```{r}
# the distances in the model compare with the original distances
plot(d, dist(points1_3))
abline(0, 1, col = "red")
```

```{r}
# mean absolute difference of the entries
diff1_3a <- mean(abs(as.matrix(dist(points1_3)) - as.matrix(d)))
diff1_3a
```

```{r}
# max absolute difference of the entries
diff1_3b <- max(abs(as.matrix(dist(points1_3)) - as.matrix(d)))
diff1_3b
```
##  Models for classic normalization data

```{r}
# 1-D model
model2_1 <- cmdscale(d1,k=1,eig=TRUE)
# points in the model
points2_1 <- model2_1$points
points2_1
```

```{r}
# eigenvalues in the model
eigs2_1 <- model2_1$eig
eigs2_1
```

```{r}
# Plot eigenvalues
plot(model2_1$eig, xlab = "Index for cities", ylab = "Eigenvalues")
```

```{r}
# GOF of the 1-D model
GOF2_1 <- model2_1$GOF
GOF2_1
```

```{r}
m2_1 = data.frame(model2_1$points, 0)
# Visulize 1-D model
textplot(m2_1[, 1], m2_1[, 2], gsub("(ˆ[ˆ\\s]+\\s{1})", "", cityIndex, perl = TRUE), xlim = c(-10, 5), ylim = c(-10, 5), cex = 0.6)
```

```{r}
# the distances in the model compare with the original distances
plot(d1, dist(points2_1))
abline(0, 1, col = "red")
```

```{r}
# mean absolute difference of the entries
diff2_1a <- mean(abs(as.matrix(dist(m2_1)) - as.matrix(d1)))
diff2_1a
```

```{r}
# max absolute difference of the entries
diff2_1b <- max(abs(as.matrix(dist(m2_1)) - as.matrix(d1)))
diff2_1b
```

```{r}
# 2-D model
model2_2 <- cmdscale(d1,k=2,eig=TRUE)
# points in the model
points2_2 <- model2_2$points
points2_2
```

```{r}
# eigenvalues in the model
eigs2_2 <- model2_2$eig
eigs2_2
```

```{r}
# Plot eigenvalues
plot(model2_2$eig, xlab = "Index for cities", ylab = "Eigenvalues")
```

```{r}
# GOF of the 2-D model
GOF2_2 <- model2_2$GOF
GOF2_2
```

```{r}
m2_2 = data.frame(model2_2$points, 0)
# Visulize 2-D model
textplot(m2_2[, 1], m2_2[, 2], gsub("(ˆ[ˆ\\s]+\\s{1})", "", cityIndex, perl = TRUE), xlim = c(-10, 5), ylim = c(-10, 5), cex = 0.6)
```

```{r}
# the distances in the model compare with the original distances
plot(d1, dist(points2_2))
abline(0, 1, col = "red")
```

```{r}
# mean absolute difference of the entries
diff2_2a <- mean(abs(as.matrix(dist(m2_2)) - as.matrix(d1)))
diff2_2a
```

```{r}
# max absolute difference of the entries
diff2_2b <- max(abs(as.matrix(dist(m2_2)) - as.matrix(d1)))
diff2_2b
```

```{r}
# 3-D model
model2_3 <- cmdscale(d1,k=3,eig=TRUE)
# points in the model
points2_3 <- model2_3$points
points2_3
```

```{r}
# eigenvalues in the model
eigs2_3 <- model2_3$eig
eigs2_3
```

```{r}
# Plot eigenvalues
plot(model2_3$eig, xlab = "Index for cities", ylab = "Eigenvalues")
```

```{r}
# GOF of the 3-D model
GOF2_3 <- model2_3$GOF
GOF2_3
```

```{r}
m2_3 = data.frame(model2_3$points, 0)
```

```{r}
# the distances in the model compare with the original distances
plot(d1, dist(points2_3))
abline(0, 1, col = "red")
```

```{r}
# mean absolute difference of the entries
diff2_3a <- mean(abs(as.matrix(dist(m2_3)) - as.matrix(d1)))
diff2_3a
```
```{r}
# max absolute difference of the entries
diff2_3b <- max(abs(as.matrix(dist(m2_3)) - as.matrix(d1)))
diff2_3b
```

##  Models for another normalization data

```{r}
# 1-D model
model3_1 <- cmdscale(d2,k=1,eig=TRUE)
# points in the model
points3_1 <- model3_1$points
points3_1
```

```{r}
# eigenvalues in the model
eigs3_1 <- model3_1$eig
eigs3_1
```

```{r}
# Plot eigenvalues
plot(model3_1$eig, xlab = "Index for cities", ylab = "Eigenvalues")
```

```{r}
# GOF of the 1-D model
GOF3_1 <- model3_1$GOF
GOF3_1
```

```{r}
m3_1 = data.frame(model3_1$points, 0)
# Visulize 1-D model
textplot(m3_1[, 1], m3_1[, 2], gsub("(ˆ[ˆ\\s]+\\s{1})", "", cityIndex, perl = TRUE), xlim = c(-2, 5), ylim = c(-5, 5), cex = 0.6)
```

```{r}
# the distances in the model compare with the original distances
plot(d2, dist(points3_1))
abline(0, 1, col = "red")
```

```{r}
# mean absolute difference of the entries
diff3_1a <- mean(abs(as.matrix(dist(m3_1)) - as.matrix(d2)))
diff3_1a
```
```{r}
# max absolute difference of the entries
diff3_1b <- max(abs(as.matrix(dist(m3_1)) - as.matrix(d2)))
diff3_1b
```

```{r}
# 2-D model
model3_2 <- cmdscale(d2,k=2,eig=TRUE)
# points in the model
points3_2 <- model3_2$points
points3_2
```

```{r}
# eigenvalues in the model
eigs3_2 <- model3_2$eig
eigs3_2
```

```{r}
# Plot eigenvalues
plot(model3_2$eig, xlab = "Index for cities", ylab = "Eigenvalues")
```

```{r}
# GOF of the 2-D model
GOF3_2 <- model3_2$GOF
GOF3_2
```

```{r}
m3_2 = data.frame(model3_2$points, 0)
# Visulize 2-D model
textplot(m3_2[, 1], m3_2[, 2], gsub("(ˆ[ˆ\\s]+\\s{1})", "", cityIndex, perl = TRUE), xlim = c(-2, 3), ylim = c(-4, 4), cex = 0.6)
```

```{r}
# the distances in the model compare with the original distances
plot(d2, dist(points3_2))
abline(0, 1, col = "red")
```

```{r}
# mean absolute difference of the entries
diff3_2a <- mean(abs(as.matrix(dist(m3_2)) - as.matrix(d2)))
diff3_2a
```

```{r}
# max absolute difference of the entries
diff3_2b <- max(abs(as.matrix(dist(m3_2)) - as.matrix(d2)))
diff3_2b
```

```{r}
# 3-D model
model3_3 <- cmdscale(d2,k=3,eig=TRUE)
# points in the model
points3_3 <- model3_3$points
points3_3
```

```{r}
# eigenvalues in the model
eigs3_3 <- model3_3$eig
eigs3_3
```

```{r}
# Plot eigenvalues
plot(model3_3$eig, xlab = "Index for cities", ylab = "Eigenvalues")
```

```{r}
# GOF of the 3-D model
GOF3_3 <- model3_3$GOF
GOF3_3
```

```{r}
m3_3 = data.frame(model3_3$points, 0)
```

```{r}
# the distances in the model compare with the original distances
plot(d2, dist(points3_3))
abline(0, 1, col = "red")
```

```{r}
# mean absolute difference of the entries
diff3_3a <- mean(abs(as.matrix(dist(m3_3)) - as.matrix(d2)))
diff3_3a
```

```{r}
# max absolute difference of the entries
diff3_3b <- max(abs(as.matrix(dist(m3_3)) - as.matrix(d2)))
diff3_3b
```

```{r}
# Compare the GOF value from 2 normalization methods we used above in 3 different dimensions.
plot(c(GOF2_1[1], GOF2_2[1], GOF2_3[1]), ylim = c(0.6, 1.0), xlab = "Dimension", ylab = "GOF", col = "blue")
points(c(GOF3_1[1], GOF3_2[1], GOF3_3[1]), col = "red")
legend("topright",c("standardized with classic method", "standardized with \"max\ and \"min\" method"), text.col = c("blue","red"), cex =
1)
```

```{r}
# Compare the mean absolute difference between the distance from the data and distance we calculated by using 2 methods above in 3 different dimensions.
plot(c(diff2_1a, diff2_2a, diff2_3a), ylim = c(0, 1), xlab = "Dimension", ylab = "Mean absolute
difference", col = "blue")
points(c(diff3_1a, diff3_2a, diff3_3a), col = "red")
legend("topright",c("standardized with classic method", "standardized with \"max\ and \"min\" method"), text.col = c("blue","red"), cex =
1)
```

```{r}
# Compare the maximum absolute difference between the distance from the data and distance we calculated by using 2 methods above in 3 different dimensions.
plot(c(diff2_1b, diff2_2b, diff2_3b), ylim = c(0, 3), xlab = "Dimension", ylab = "Maximum absolute
difference", col = "blue")
points(c(diff3_1b, diff3_2b, diff3_3b), col = "red")
legend("topright",c("standardized with classic method", "standardized with \"max\ and \"min\" method"), text.col = c("blue","red"), cex =
1)
```

```{r}
# Calculate the correlation value for different methods in different dimensions.
correlation2_3x <- cor(points2_3[, 1], normalize1)
correlation3_3x <- cor(points3_3[, 1], normalize2)
correlation2_3y <- cor(points2_3[, 2], normalize1)
correlation3_3y <- cor(points3_3[, 2], normalize2)
correlation2_3z <- cor(points2_3[, 3], normalize1)
correlation3_3z <- cor(points3_3[, 3], normalize2)
```